---
title: "Loading spatial data in R"
author: "Matt Williamson"
date: "9/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## The data

One of the other datasets we'll use fairly regulrarly during this class is the United States Protected Areas Database ([PAD-US](https://www.usgs.gov/core-science-systems/science-analytics-and-synthesis/gap/science/protected-areas)). PAD-US is America’s official national inventory of U.S. terrestrial and marine protected areas that are dedicated to the preservation of biological diversity and to other natural, recreation and cultural uses, managed for these purposes through legal or other effective means. The PADUS provides a lot of interesting information on land tenure arrangements in the US. Ecologists often rely on comparisons to these protected ares to understand the effects of anthropogenic change. Finally, you'll probably have to make a map at some point in your life and these protected areas are often useful tools for helping 'orient' people to the landscape you are mapping. As an additional bonus, the geometries in this dataset can be a real challenge to work with so you'll get a chance to practice some important diagnostics and trouble-shooting.

Speaking of land tenure, the raster dataset we'll use this week comes from the [PLACES lab](https://placeslab.org/places/) at Boston University. This data depict land values across the contiguous United States at a fairly high resolution. The PLACES data was developed to better understand the costs, benefits, and motivaitons for private land conservation. The development and validation of this data is described in this [article](https://www.pnas.org/content/117/47/29577).


## Your assignment

The goal for this assignment is to ensure that you:
- can load different spatial data into R using different packages

- can access the different components of spatial data within R

- can verify projection information and modify to ensure data aligns

- can make simple plots to 'gut check' your data


## Step 1: Load your packages
You'll need the `raster`, `terra`, `sp`, `rgdal`, `sf`, and `tidyverse` packages to complete this assignment. Load them in the code chunk below.

```{r packageload}

```


## Step 2: Load polygon data

The data for this lab are in our shared folder (located at `opt/data/session04/`). There are several shapefiles in that folder (denoted by the `.shp` suffix in the file name). Use our approach for finding all of the files that match a given pattern (remember last week) to load all of the shapefiles. You should do this both with `st_read` and `readOGR`. Make sure you assign them to different objects. Add your code to the code chunk below:

```{r loadpoldata}

```
__Question 1__: How do the objects created by `st_read` and `readOGR` compare (remember that you can use `[[` to access the individual objects produced by lapply)?


## Step 3: Are the geometries valid?
Now we have to make sure the shapefiles have valid geometries. Check them and, where necessary, fix them in the code chunk below.
```{r fixgeom}

```



## Step 4: Load the raster data

Following the same approach you used above, load all of the raster data from our shared folder. You should use `raster::raster` and `terra::rast` so that you can see how the objects differ when you load them.

```{r loadrastdata}

```

## Step 5: Compare the state and PA polygons

Now that you have lists containing all of the shapefiles in our class folder, lets get the two we'll use for the rest of the exercise. Using `[[`, assign the `regionalstates.shp`, `regionalPAs1.shp`,  `regionalPAs2.shp`, and  `regionalPAs3.shp` to their own objects. Do the same with the `Regval.tif`. Now, let's make sure they have the same projections by comparing the CRS of each object. You can use any of the ways we discussed in class. Do that in the code chunk below.

```{r compareproj}

```

__Question 2__: Do all the projections match? How might you go about choosing which projection you want to use going forward? _Bonus_ one of the versions of the PA data is missing a projection entirely, how can you fix that? Insert a code chunk and demonstrate.



## Step 6: Project the data

Before we can do much with the different datasets, we've got to get them into the same projections so they'll line up. Use the code chunk below to get the `regionalstates.shp`, one of the protected area shapefiles, and the land value raster into the same projection.

```{r projectdata}

```

## Step 7: Make a simple plot

Now that the data is all aligned you should be able to plot them one on top of the other. This is a good way to double check that all of your projections worked! Develop the plot in the code chunk below. Remember that plots build like layers in a cake, so you'll want to have the raster as the base, then the states, then the protected areas.

```{r plotdata}

````

## Bonus
The `regionalcessions.shp` file depicts the various land cessions that form the basis of the Land Grab University analysis. Follow the steps above to add the land cessions to the land value and protected area map. Show your code in appropriate code chunks, then offer any thoughts you may have about the spatial patterns that you see in your plot.

