---
title: "Loading spatial data in R"
author: "Matt Williamson"
date: "9/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## The data

One of the other datasets we'll use fairly regulrarly during this class is the United States Protected Areas Database ([PAD-US](https://www.usgs.gov/core-science-systems/science-analytics-and-synthesis/gap/science/protected-areas)). PAD-US is America’s official national inventory of U.S. terrestrial and marine protected areas that are dedicated to the preservation of biological diversity and to other natural, recreation and cultural uses, managed for these purposes through legal or other effective means. The PADUS provides a lot of interesting information on land tenure arrangements in the US. Ecologists often rely on comparisons to these protected ares to understand the effects of anthropogenic change. Finally, you'll probably have to make a map at some point in your life and these protected areas are often useful tools for helping 'orient' people to the landscape you are mapping. As an additional bonus, the geometries in this dataset can be a real challenge to work with so you'll get a chance to practice some important diagnostics and trouble-shooting.

Speaking of land tenure, the raster dataset we'll use this week comes from the [PLACES lab](https://placeslab.org/places/) at Boston University. This data depict land values across the contiguous United States at a fairly high resolution. The PLACES data was developed to better understand the costs, benefits, and motivaitons for private land conservation. The development and validation of this data is described in this [article](https://www.pnas.org/content/117/47/29577).


## Your assignment

The goal for this assignment is to ensure that you:
- can load different spatial data into R using different packages

- can access the different components of spatial data within R

- can verify projection information and modify to ensure data aligns

- can make simple plots to 'gut check' your data
- 

## Step 1: Load your packages
You'll need the `raster`, `terra`, `sp`, `rgdal`, `sf`, and `tidyverse` packages to complete this assignment. Load them in the code chunk below.

```{r packageload}

```


## Step 2: Load polygon data

The data for this lab are in our shared folder (located at `opt/data/session04/`). There are several shapefiles in that folder (denoted by the `.shp` suffix in the file name). Use our approach for finding all of the files that match a given pattern (remember last week) to load all of the shapefiles. You should do this both with `st_read` and `readOGR`. Make sure you assign them to different objects. Add your code to the code chunk below:

```{r loaddata}

```
__Question 1__: How do the objects created by `st_read` and `readOGR` compare (remember that you can use `[[` to access the individual objects produced by lapply)?

## Step 3: Compare the state and PA polygons


